---
title: "Report on Linear Programming"
author: "Himanshu Poddar"
date: "29 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Programming

###Firstly Let us understand What is Linear Programming?

Linear Programming is a mathematical technique for maximizing or minimizing a linear function of several variables, such as cost.
One of the reason for the popularity of LP is that it allows to model a large variety of situations with a simple framework.

#Linear Programming in our daily analytics operations.

Linear programming is used for obtaining the most optimal solution for a problem with given constraints. In linear programming, we formulate our real life problem into a mathematical model. It involves an objective function, linear inequalities with subject to constraints.
LPP is all about optimising(that is, either minimise or maximise) the value of a linear objective function of a vector of decision variables, considering that the variables can only take the values defined by a set of linear constraints. LP is a case of mathematical programming where objective function and constraints are linear.

Example of a linear programming problem
Let's say a delivery man has 6 packages to deliver in a day. The warehouse is located at point A. The 6 delivery destinations are given by U, V, W, X, Y and Z. The numbers on the lines indicate the distance between the cities. To save on fuel and time the delivery person wants to take the shortest route.
So, the delivery person will calculate different routes for going to all the 6 destinations and then come up with the shortest route. This technique of choosing the shortest route is called linear programming.
 
 For solving a problem on LPP we need to formulate the problem
 
 eg:
 1. The statement which we have to either maximise or minimise (denoted by Z).
 MAX/MIN z = c1x1 + c2x2 + ...... + cNxN

2. Given constraints
We need to consider the constraints given in the problem statement.

A1x1 + B1x2 + ............ + Z1xN <= C1

A2x1 + B2x2 + ............ + Z2xN <= C2

                 where A1,A2,B1.... are constants.
                 and C1,C2 are upper or lower bounds.
                 
We have to consider all the constraint to either maximise or minimise Z.


#Solving LPP problems with R

In R we will use a package called lpSolve which is a package used to solve LPP.
`r library(lpSolve)`


There are different models present in the lp package .
eg : "lp" ,"lp.assign" , "lp.transport" , "make.q8"

We will be using lp to solve using Linear Programming Mode.


###EXAMPLE
A car company produces 2 models, model A and model B. Long-term projections indicate an expected demand of at least 100 model A cars and 80 model B cars each day. Because of limitations on production capacity, no more than 200 model A cars and 170 model B cars can be made daily. To satisfy a shipping contract, a total of at least 200 cars much be shipped each day. If each model A car sold results in a $2000 loss, but each model B car produces a $5000 profit, how many of each type should be made daily to maximize net profits?

##Solution:

Let us consider
X= number of model A car Y=number of model B car

Since $2000 is the loss, we take it as negative

-2000x+5000y=Z

We have to maximise Z

And the constraints are

Expected demand of model A      : x>=100

Expected demand of model B      : y>=80 

Limit of production of model A  : x<=200

Limit of production of model B  : y<=170

Shipping contract               : x+y>=200

The linear equations can be written as

1x + 0y >= 100

0x + 1y >= 80

1x +0y <= 200

0x + 1y <= 170

1x + 1y >= 200

therefore,the constraint matrix formed would be the coefficient of these equations,i.e,

                     | 1    0 |
                     | 0    1 |
                     | 1    0 |
                     | 0    1 |
                     | 1    1 |
                               {5X2}
                              
The inequality array or the direction of the constraint can be written as,
                    
                    [ ">=" , ">=" , "<=" , "<=" , ">=" ]
                    
The array for the RHS value of the constraint array is given as,
                    
                    [ 100 , 80 , 200 , 170 , 200 ]
                               
                               
#Code

```{r}
#Creating the constraint matrix
l.constr <- matrix(c(1,0,0,1,1,0,0,1,1,1),nrow = 5,byrow = TRUE)
 #Objective function
l.obj <- c(-2000,5000)
#Inequality of constraints
l.dir <- c(">=",">=","<=", "<=",">=")
#RHS of constraints
l.rhs <- c(100,80,200,170,200)
#Passing the values to our LPP function
lps <- lp ("max", l.obj, l.constr, l.dir, l.rhs,compute.sens = TRUE)
#Printing the solution
print(lps$solution)
```



##Conclusion
Therefore it can be seen that we have to produce 100 units of Model A and 170 of Model B on daily basis to maximize net profits.